---
title: 'ANOVA and T-test R Scripts'
author: "Abraham Alex"
output: 
  html_document:
    toc: false
    toc_float: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


# 2 (factors of validity) X 2 (time of rating) X 2 (face gender) Within-Subject ANOVA

```{r}
mydata = read.csv("Ratings_ANOVA.csv")
#mydata$Subject = factor(mydata$Subject)

listOfRatings = c()
for(x in 1:50) {
  listOfRatings = append(listOfRatings, mydata[x, c(1:8)]) }
listOfRatings = as.numeric(listOfRatings)

# 2 X 2 X 2 ANOVA with the following independent variables and levels
#   factors of validity: I means incongruent and C means congruent
#   face gender: F means female and M means male
#   time of rating: B means beginning and E means end of rating

mydata_revised = data.frame(subject = c(rep(c(1:50), each = 8)), 
                            factors_of_validity = c(rep(c("I","I","I","I","C","C","C","C"), times = 50)),
                            face_gender = c(rep(c("F","M"), times = 200)), 
                            time_of_rating = c(rep(c("B","B","E","E"), times = 100)), 
                            ratings = listOfRatings)

mydata_revised$subject = factor(mydata_revised$subject)
#print(mydata_revised)

#aov.bww <- aov(y ~ b1*w1*w2 + Error(subject/(w1*w2)) + b1, data=data_long)
aov.bww = aov(ratings ~ factors_of_validity*face_gender*time_of_rating + Error(subject/(factors_of_validity*face_gender*time_of_rating)), data=mydata_revised)
summary(aov.bww)
```

# T-test Comparison of Factors of Validity at the Initial Time of Rating

```{r}
beginningCongruentRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "B"),    which(mydata_revised$factors_of_validity == "C" ))]
beginningIncongruentRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "B"), which(mydata_revised$factors_of_validity == "I" ))]

t.test(beginningCongruentRatings, beginningIncongruentRatings, var.equal = T)
```

# T-test Comparison of Face Gender at the Initial Time of Rating

```{r}
beginningFemaleRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "B"), which(mydata_revised$face_gender == "F" ))]
beginningMaleRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "B"), which(mydata_revised$face_gender == "M" ))]

t.test(beginningFemaleRatings, beginningMaleRatings, var.equal = T)
```

# T-test Comparison of Face Gender at the Final Time of Rating

```{r}
endingFemaleRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "E"), which(mydata_revised$face_gender == "F" ))]
endingMaleRatings = mydata_revised$ratings[intersect(which(mydata_revised$time_of_rating == "E"), which(mydata_revised$face_gender == "M" ))]

t.test(endingFemaleRatings, endingMaleRatings, var.equal = T)
```

# 2 (Factors of Validity) X 2 (Face Gender) ANOVA at the Initial Time of Rating

```{r}
initialRatings = mydata_revised$ratings[which(mydata_revised$time_of_rating == "B")]
initialFactorsOfValidity = mydata_revised$factors_of_validity[which(mydata_revised$time_of_rating == "B")]
initialFaceGender = mydata_revised$face_gender[which(mydata_revised$time_of_rating == "B")]
initialSubject = mydata_revised$subject[which(mydata_revised$time_of_rating == "B")]


aov.bww = aov(initialRatings ~ initialFactorsOfValidity*initialFaceGender + Error(initialSubject/(initialFactorsOfValidity*initialFaceGender)))

summary(aov.bww)
```


# 2 (Factors of Validity) X 2 (Face Gender) ANOVA at the Final Time of Rating

```{r}
finalRatings = mydata_revised$ratings[which(mydata_revised$time_of_rating == "E")]
finalFactorsOfValidity = mydata_revised$factors_of_validity[which(mydata_revised$time_of_rating == "E")]

finalFaceGender = mydata_revised$face_gender[which(mydata_revised$time_of_rating == "E")]
finalSubject = mydata_revised$subject[which(mydata_revised$time_of_rating == "E")]


aov.bww = aov(finalRatings ~ finalFactorsOfValidity*finalFaceGender + Error(finalSubject/(finalFactorsOfValidity*finalFaceGender)))

summary(aov.bww)
```

